---
output:
  md_document:
    variant: markdown_github
---

<!-- README.md is generated from README.Rmd. Please edit that file -->

```{r, include = FALSE}
devtools::load_all('~/ignore/knitLatex')
knitr_sethooks()
knitr::opts_chunk$set(
  collapse = TRUE,
  fig.path = "README-"
)
```


This package includes several helper functions for working with knitr and latex.
It includes xTab for creating traditional latex tables, lTab for generating
Longtable environments, and sTab for generating a Supertabular environment.
Additionally, this package contains a knit_sethooks function. knitr_sethooks
serves two purposes. First, it fixes a well-known bug which distorts the
"results='asis'" command when used in conjunction with userdifiend commands.
Second, it provides a com command (\<\<com=TRUE\>\>=) which renders the output
from knitr as a latex command.

In the examples below, the code is shown as it would be provided in a knitr
chunk, with the resultant latex. If this was an actual latex file, setting
results = 'asis' would eliminate the comment markings (which is generally what
you would want). Additionally, the examples below show only the basic usages of
each function. For a more detailed examination, you should view the vigenttes
(provided in inst/doc if you are viewing on github). To see a complete working
example, see the documentation in inst/example (again if you are viewing on
github).

# xTab

```{r}
xTab(mtcars)
```

# sTab

```{r}
sTab(mtcars)
```


# lTab

```{r}
lTab(mtcars)
```
