---
title: "xTab"
author: "John Shea"
date: "`r Sys.Date()`"
output:
  rmarkdown::html_vignette:
    toc: true
vignette: >
  %\VignetteIndexEntry{xTab}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

xTab is a command the produces a standard latex table/tabular environment

```{r, echo = FALSE}
knitr::opts_chunk$set(collapse = TRUE)
cars  <- mtcars[1:10,1:5]
```

For all of the examples that follow, we will be using the first team rows and
five columns of the mtcars data.frame, which we have saved in the variable
'cars'. It is improtant to note that, when passing latex commands to an xTab
argument, backslashes must be escaped. For example, you need to pass '\\\\hline'
to produce \\hline in the latex document. Similarly you must pass'\\\\\\\\' to
produce \\\\.

## Basics

### Standard xTab table

To produce a latex table, simply pass a matrix or a data.frame to the xTab
function.

```{r}
xTab(cars)
```

### Labels

```{r}
xTab(cars, label = 'tab:mytable')
```

### Captions

Place a caption at the top of the table.

```{r}
xTab(cars, caption.top = 'my table')
```

Place a caption at the bottom of the table

```{r}
xTab(cars, caption.bottom = 'my table')
```

### Booktabs

Setting booktabs = TRUE sets the toprule, midrule, and bottomrule arguments to
\\toprule, \\midrule, and \\bottomrule respectively. If any of those options are
explicitly set, the booktabs value has no effect. When using booktabs rules,
regardless of whether you set booktabs = TRUE or set them individually, make
sure to include \\usepackage(booktabs) in your latex document.

```{r}
xTab(cars, booktabs = TRUE)
```

```{r}
xTab(cars, booktabs = TRUE, midrule = '\\hline')
```
## Headers

By default, xTab will use the column names for the head of the table(as
demonstrated in the above examples). To customize the table head simply pass in
the appropriate latex to the 'head' argument. The values of toprule and midrule
are still used and should not be set in the head argument. If you do not want
them included in your custom header, set either or both to NULL

```{r}
xTab(cars,
     head = 'col1 & col 2 & col3 & \\eta & col5 \\\\')
```

Pass in NULL to avoid a table head. In the case of a NULL head, toprule and
midrule will not be used.

```{r}
xTab(cars, head = NULL)
```

## Rows

### Rownames

When including row names in a table, by default xTab will use an empty
columanname for the 'rownames' column.

```{r}
xTab(cars, rows = TRUE)
```

### Rownames with custom header

When providing a custom head with rows set to TRUE, remember to account for the
extra column produced by the row names

```{r}
xTab(cars,
     rows =  TRUE,
     head = 'rows & col1 & col2 & col3 & \\eta & col5 \\\\')
```

### Row separator

Any arbitrary latex command can be inserted between each row, but the mostcommon
are \\hline and \\midrule. To use \\midrule, \\usepackage(booktabs) must be
declared in the preamble of the latex document, but booktabs = TRUE does NOT
need to be set on the table.

```{r}
xTab(cars, rowsep = '\\hline')
```

```{r}
xTab(cars, rowsep = '\\midrule')
```

### Include a vertical bar between each column

```{r}
xTab(cars, colsep = '|')
```
