---
title: "xTab"
author: "John Shea"
date: "`r Sys.Date()`"
output:
  rmarkdown::html_vignette:
    toc: true
vignette: >
  %\VignetteIndexEntry{xTab}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

xTab is a command the produces a standard latex table/tabular environment

```{r, echo = FALSE}
knitr::opts_chunk$set(collapse = TRUE)
cars  <- mtcars[1:10,1:5]
```

For all of the examples that follow, we will be using the first team rows and
five columns of the mtcars data.frame, which we have saved in the variable
'cars'

## Standard xTab table

To produce a latex table, simply pass a matrix or a data.frame to the xTab
function.

```{r}
xTab(cars)
```

## Include a label

```{r}
xTab(cars, label = 'tab:mytable')
```

## Place a caption at the top of the table

```{r}
xTab(cars, caption.top = 'my table')
```

## Place a caption at the bottom of the table
```{r}
xTab(cars, caption.bottom = 'my table')
```
## Customize table head

By default, xTab will use the column names for the head of the table. To
customize the table head simply pass in the appropriate latex to the 'head'
argument. Pass in NULL to avoid a table head.

```{r}
xTab(cars,
     head = 'col1 & col 2 & col3 & \\eta & col5 \\\\')
```

## Produce a table without a table head

```{r}
xTab(cars, head = NULL)
```

## Include rownames in an xTab table

When including row names in a table, by default xTab will use an empty
columanname for the 'rownames' column.

```{r}
xTab(cars, rows = TRUE)
```

## Include rownames and a custom head

When providing a custom head with rows set to TRUE, remember to account for the
extra column produced by the row names

```{r}
xTab(cars,
     rows =  TRUE,
     head = 'rows & col1 & col2 & col3 & \\eta & col5 \\\\')
```

## Use booktabs rules instead of default \\hline

```{r}
xTab(cars, booktabs = TRUE)
```

## Separate each row with a latex command

Any arbitrary latex command can be inserted between each row, but the mostcommon
are \\hline and \\midrule. To use \\midrule, \\usepackage(booktabs) must be
declared in the preamble of the latex document, but booktabs = TRUE does NOT
need to be set on the table.

```{r}
xTab(cars, rowsep = '\\hline')
```

```{r}
xTab(cars, rowsep = '\\midrule')
```

## Include a vertical bar between each column

```{r}
xTab(cars, colsep = '|')
```
