---
title: "xTab"
author: "John Shea"
date: "`r Sys.Date()`"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{xTab}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

# xTab is a command the produces a standard latex table/tabular environment

```{r, echo = FALSE}
knitr::opts_chunk$set(collapse = TRUE)
cars  <- cars[1:10,]
```

For all of the examples taht follow, we will be using the first team rows of the
mtcars data.frame, which we have saved in the variable 'cars'

## standard xTab table

To produce a latex table, simply pass a matrix or a data.frame to the xTab
function.

```{r}
xTab(cars)
```

## include a label for the table

```{r}
xTab(cars, label = 'tab:mytable')
```

## place a caption at the top of the table

```{r}
xTab(cars, caption.top = 'my table')
```

## place a caption at the bottom of the table
```{r}
xTab(cars, caption.bottom = 'my table')
```
## customize table head

By default, xTab will use the column names for the head of the table. To
customize the table head simply pass in the appropriate latex to the 'head'
argument. Pass in NULL to avoid a table head.

```{r}
xTab(cars,
     head = 'col1 & col 2 & col3 & \\eta & col5 \\\\')
```

## include rownames in an xTab table
```{r}
xTab(cars, rows = TRUE)
```

## include rownames and a custom head

When providing a custom head with rows set to TRUE, remember to account for the
extra column produced by the row names

```{r}
xTab(cars,
     rows =  TRUE,
     head = 'col1 & col2 & col3 & \\eta & col5 \\\\')
```

## separate each row with an \\hline
```{r}
xTab(cars, rowsep = '\\hline')
```

## include a vertical bar between each column
```{r}
xTab(cars, colsep = '|')
```
